\textbf{Are regime changes always bad economics? Evidence from daily financial data}

### 2021-11-17 doc/archive/paper.tex
Additional changes to note manuscript and organization
- Minor changes to abstract, theory, and conclusion.
- Created anonymous note manuscript.
- Moved old .tex files to archive folder.

\setcounter{tocdepth}{1}
\begin{document}
%\author{Devin Incerti\thanks{Senior Data Scientist, Genentech. devin.incerti@gmail.com} and Trevor Incerti\thanks{Corresponding author. PhD Candidate in the Department of Political Science, Yale University. trevor.incerti@yale.edu.}}
\title{\textbf{Are regime changes always bad for markets? Evidence from daily financial data}} 
\date{\today}
\maketitle
\thispagestyle{empty}

@@

We first attempt to explore these mechanisms by aggregating all of the events in our sample by whether they resulted in a shift in an authoritarian or democratic direction.\footnote{As defined by the Polity project.} We find suggestive evidence that regime changes associated with authoritarian shifts are on average perceived negatively by investors (see \autoref{tab:AR-auth}), while those that move governance in a democratic direction are on average perceived positively (see \autoref{tab:AR-dem}). We refer to this evidence as suggestive despite statistical significance due to the small sample of cases which fit these criteria, particularly with regard to democratic shifts. We observe ten cases associated with an authoritarian shift in governance. Only two of these events result in positive CARs, and neither are significantly different from zero. Five cases resulted in a democratic shift in governance. Three of these five cases result in positive CARs. Of the two negative CARs, one is not significantly different from zero, and the other is associated with a forced market closure that lasted 17 days. However, the majority of the positive returns from democratic regime changes come from the 12\% positive CAR associated with the resignation of Ferdinand Marcos in the Philippines in 1986.

>>>>

\setcounter{tocdepth}{1}
\begin{document}
\author{Devin Incerti\thanks{Senior Data Scientist, Genentech. devin.incerti@gmail.com} and Trevor Incerti\thanks{Corresponding author. PhD Candidate in the Department of Political Science, Yale University. trevor.incerti@yale.edu.}}
\title{\textbf{Are regime changes always bad economics? Evidence from daily financial data \thanks{We are grateful to Peter Aronow, Matthew Graham, and Frances Rosenbluth for helpful feedback and suggestions. Authors are listed in alphabetic order, implying equal authorship. Any and all errors are our own.}}}
\date{\today}
\maketitle
\thispagestyle{empty}

@@

We first attempt to explore these mechanisms by aggregating all of the events in our sample by whether they resulted in a shift in an authoritarian or democratic direction.\footnote{As defined by the Polity project} We find suggestive evidence that regime changes associated with authoritarian shifts are on average perceived negatively by investors (see \autoref{tab:AR-auth}), while those that move governance in a democratic direction are on average perceived positively (see \autoref{tab:AR-dem}). We refer to this evidence as suggestive despite statistical significance due to the small sample of cases which fit these criteria, particularly with regard to democratic shifts. We observe ten cases associated with an authoritarian shift in governance. Only two of these events result in positive CARs, and neither are significantly different from zero. Five cases resulted in a democratic shift in governance. Three of these five cases result in positive CARs. Of the two negative CARs, one is not significantly different from zero, and the other is associated with a forced market closure that lasted 17 days. However, the majority of the positive returns from democratic regime changes come from the 12\% positive CAR associated with the resignation of Ferdinand Marcos in the Philippines in 1986.

### 2021-11-17 doc/regime_changes_note_anonymous.tex
Additional changes to note manuscript and organization
- Minor changes to abstract, theory, and conclusion.
- Created anonymous note manuscript.
- Moved old .tex files to archive folder.

\setcounter{tocdepth}{1}
\begin{document}
\author{Devin Incerti\thanks{Principal Data Scientist, Genentech. devin.incerti@gmail.com} and Trevor Incerti\thanks{Corresponding author. PhD Candidate in the Department of Political Science, Yale University. trevor.incerti@yale.edu.}}
\title{\textbf{Are regime changes always bad economics? Evidence from daily financial data \thanks{We are grateful to P.M. Aronow, Matthew Graham, Frances Rosenbluth, and Hikaru Yamagishi for helpful feedback and suggestions. Authors are listed in alphabetic order, implying equal authorship. Any and all errors are our own.}}}
\date{\today}
\maketitle
\thispagestyle{empty}

\singlespacing
\begin{abstract}
\noindent
Political instability is commonly thought to discourage investment and reduce economic growth. We challenge this consensus by showing that instability does not systematically depress access to financial capital. Using an event study approach, we examine daily returns of national stock indices in every country that experienced an irregular regime change subject to data availability. Returns following resignations are large and positive (+4\%), while those following assassinations are negative and smaller in magnitude (-2\%). The impact of coups tends to be negative (-2\%), but we show that a pro-capitalist coup results in large positive returns (+10\%). We find suggestive evidence that authoritarian or anti-capitalist regime changes are more likely to lead to capital flight than democratic or pro-business changes.  The immediate impact of political instability on capital access is therefore dependent on the type of regime change and its expected impact on economic policy.

@@

Conventional theory and cross-country empirical evidence suggests that political instability increases country risk and therefore depresses financial returns and economic growth \citep{irshad2017relationship, le2006political, lensink2000capital, boutchkova2012precarious, lehkonen2015democracy}, and increases volatility \citep{irshad2017relationship, bialkowski2008stock, leblang2005government, jensen2005market, liu2015economic}. But if capital flows reflect investor expectations about future economic growth and returns, such hypotheses may be too simplistic. For example, capital access may increase in response to a coup if the current regime is anti-business or anti-global.

@@

Debate also exists on ``good coups''---i.e., coups that lead to democratization or economic liberalization---and their effect on economic growth.  Most research suggests these are not the norm \citep{derpanopoulos2015coups, powell2011global, thyne2016coup, varol2011democratic}. Others argue the negative effects of uncertainty dominate any positive effects \citep{alesina1996political}, or that the expected impact of coups is neutral as some enhance growth while others depress it \citep{londregan1990poverty}. But while ``good coups'' may be the minority, they have become more frequent \citep{marinov2014coups} and could have positive economic effects \citep{meyersson2016political, girardi2018institution}.

@@

Previous research therefore suggests coups should on average cause capital flight, but lead to positive flows if the coup's instigators are more pro-market than the regime they replace.  Assassinations should have a neutral or negative impact as they increase uncertainty, but institutional responses to assassinations vary by country. By contrast, resignations may be viewed positively on average, as they often signal the departure of an ineffective leader.

@@

This interrupted time series approach implies that the ``control group'' being compared against is not, for example, regular regime changes or failed regime changes, but the expected returns in the same country on the same day in a but-for world in which no regime change occurred. We report abnormal returns separately for coups, assassinations and resignations. However, each event CAR is by itself also a valid causal estimate of the effect of a specific regime change.

@@

We therefore aggregate the events in our sample by whether they resulted in an authoritarian or democratic shift.\footnote{As defined by the Polity project.} We find suggestive evidence that regime changes associated with authoritarian shifts are on average perceived negatively by investors (see \autoref{tab:AR-auth}), while democratic shifts are perceived positively (see \autoref{tab:AR-dem}). Of ten authoritarian shifts, only two result in positive CARs, and neither are significantly different from zero. Three of five cases involving democratic shifts result in positive CARs, and the positive CARs persist in the follows weeks. Of the two negative CARs, one is not significantly different from zero, and the other is associated with a forced market closure that lasted 17 days. While this evidence is suggestive, it comports with previous research showing that political risk to markets declines with higher levels of democracy \citep{lehkonen2015democracy}.

@@

Conventional theory suggests political instability causes capital flight. We show that this is not always the case. Unexpected changes in ruler virtually always increase volatility, but when political instability is broken down into types, evidence emerges that some types of instability increase capital access while other types lead to capital flight. Domestic capital access can be given a boost when a new regime is expected to offer a more democratic or pro-business environment.

@@

We show that resignations increase capital access on average, but assassinations and coups tend to cause capital flight. We also find evidence that capital markets tend to prefer democratic regime changes to authoritarian shifts, but that even democratically subversive coups can boost capital access if the instigators are clearly pro-business.

>>>>

\setcounter{tocdepth}{1}
\begin{document}
%\author{Devin Incerti\thanks{Principal Data Scientist, Genentech. devin.incerti@gmail.com} and Trevor Incerti\thanks{Corresponding author. PhD Candidate in the Department of Political Science, Yale University. trevor.incerti@yale.edu.}}
\title{\textbf{Are regime changes always bad economics? Evidence from daily financial data}}
\date{\today}
\maketitle
\thispagestyle{empty}

\singlespacing
\begin{abstract}
\noindent
Political instability is commonly thought to discourage investment and reduce economic growth. We challenge this consensus by showing that instability does not systematically depress access to financial capital. Using an event study approach, we examine daily returns of national financial indices in every country that experienced an irregular regime change subject to data availability. Returns following resignations are large and positive (+4\%), while those following assassinations are negative and smaller in magnitude (-2\%). The impact of coups tends to be negative (-2\%), but we show that a pro-capitalist coup results in large positive returns (+10\%). We also find evidence that authoritarian or anti-capitalist regime changes are more likely to lead to capital flight than democratic or pro-business changes.  The immediate impact of political instability on capital access is therefore dependent on the type of regime change and its expected impact on future growth.

@@

Conventional theory and cross-country empirical evidence suggests that political instability increases country risk and therefore depresses financial returns and economic growth \citep{irshad2017relationship, le2006political, lensink2000capital, boutchkova2012precarious, lehkonen2015democracy}, and increases volatility \citep{irshad2017relationship, bialkowski2008stock, leblang2005government, jensen2005market, liu2015economic}. But if capital flows reflect expectations of future economic growth and returns, such hypotheses may be too simplistic. For example, capital access may increase in response to a coup if the current regime is anti-business or anti-global.

@@

Debate also exists on coups that lead to democratization or economic liberalization and their effect on economic growth.  Most research suggests these are not the norm \citep{derpanopoulos2015coups, powell2011global, thyne2016coup, varol2011democratic}. Others argue the negative effects of uncertainty dominate any positive effects \citep{alesina1996political}, or that the impact of coups is neutral as some enhance growth while others depress it \citep{londregan1990poverty}. But while these ``good coups'' may be the minority, they have become more frequent \citep{marinov2014coups} and could have positive economic effects \citep{meyersson2016political, girardi2018institution}. Regular regime changes such as elections have been found to boost markets following opposition victories in weak democracies \citep{pantzalis2000political}, and coups may exhibit the same pattern.

@@

Previous research therefore suggests coups may on average cause capital flight, but lead to positive flows if the coup's instigators are more democratic or pro-market than the regime they replace.  Assassinations should have a neutral or negative impact as they increase uncertainty, but institutional responses to assassinations vary by country. By contrast, resignations may be viewed positively on average, as they often signal the departure of an ineffective leader.

@@

This interrupted time series approach implies that the ``control group'' being compared against is not, for example, regular regime\footnote{Past studies have examined market reactions to regular regime changes. This work finds that volatility increases around elections---particularly close elections and those that lead to a change in government political orientation \citep{bialkowski2008stock} ---and that positive returns tend to precede elections \citep{pantzalis2000political}.} changes or failed regime changes, but the expected returns in the same country on the same day in a but-for world in which no regime change occurred. We report abnormal returns separately for coups, assassinations and resignations. However, each event CAR is by itself also a valid causal estimate of the effect of a specific regime change.

@@

We therefore aggregate the events in our sample by whether they resulted in an authoritarian or democratic shift.\footnote{As defined by the Polity project.} We find suggestive evidence that regime changes associated with authoritarian shifts are on average perceived negatively by investors (see \autoref{tab:AR-auth}), while democratic shifts are perceived positively (see \autoref{tab:AR-dem}). Of ten authoritarian shifts, only two result in positive CARs, and neither are significantly different from zero. Three of five cases involving democratic shifts result in positive CARs, and the positive CARs persist in the follows weeks. Of the two negative CARs, one is not significantly different from zero, and the other is associated with a forced market closure that lasted 17 days. While this evidence is suggestive, it comports with previous research showing that political risk to markets declines with higher levels of democracy \citep{lehkonen2015democracy}, and that markets tend to react favorably to opposition victories in weak democracies \citep{pantzalis2000political}.

@@

Conventional theory suggests political instability causes capital flight. We show that this is not always the case. Unexpected changes in ruler virtually always increase volatility, but when political instability is broken down into types, evidence emerges that some types of instability increase capital access while other types lead to capital flight.

@@

We examine changes in stock market returns surrounding politically unstable domestic events to show that resignations increase capital access on average, but assassinations and coups usually cause capital flight. We also find evidence that capital markets tend to prefer democratic regime changes to authoritarian shifts, but that even democratically subversive coups can boost capital access if the instigators are clearly pro-business.
